Dusty Days and a Vibrio Craze: The Influence of Saharan Dust Deposition on Microbial Communities in High Nutrient Coastal Waters 

### 2023-05-01 code/wrangling_code/wrangling_README.Rmd
Updated README and citations

The folder contains various cleaning codes for data sets on environmental parameters (dust, temperature, salinity), and total Vibrio counts in sea surface waters.

>>>>

The folder contains various cleaning codes:

\
(1) `dust.qmd`: This code takes dust concentration data from NRL and cleans it. Steps include, switching to the correct timezone, creating concentration sums per day, as well as combining the dust data set with the *VIbrio* enumeration data set.

\
(2) `total_copies_calculator.qmd`: This script is for cleaning qPCR results on Total Vibrio enumeration. It takes Cq value outputs from the CFX (qPCR machine) and computes it to copies of gene targets of Vibrio. The data set contains a daily and monthly time series. We will be examining daily.

\
(3) `df_master.qmd`: This script combines the three clean data sets for dust concentration, copies per mL, and environmental parameters

### 2023-05-01 data/processed_data/processed_data_README.Rmd
Updated README and citations

---

This folder contains processed data that has been cleaned from code in the `wrangling` folder.\
(1) `total_clean_copies`: Vibrio enumeration (as copies per mL) across time series.\
(2) `dust_clean.rds`: Cleaned dust data
(3) `ysi_2.csv`: Environmental parameter data
(4) `tx_master.csv` which is a compilation of all relevant data. 
(5)`dust_master.csv`: Manually created dust data examining different time points, sums, avereages (helpful for statistics)

>>>>

------------------------------------------------------------------------

This folder contains processed data that has been cleaned from code in the `wrangling` folder.\

## `total_clean_copies.csv`

Vibrio enumeration as gene copies per mL across time series

#### Variables

date, site, type (monthly or daily time series), and copies_mL.

## `dust_clean.rds`

Cleaned dust data

#### Variables

date, SUM (AOT summed by date), hr_cst (hour of collection in CST), and dust (AOT)

## `ysi_2.csv`

Environmental parameter data-- did not need cleaning

#### Variables

site, date, chlorophyll (chl), salinity (sal), temperature (temp), Dissolved Oxygen in mg/L (do_mgl), Dissolved Oxygen in % (do_per), pH, secchi depth (secchi), ammonium (amm), Nitrite+Nitrate (nn), orthophosphate (orthop), silicate (sil), dissolved inorganic nitrogen (din), dissolved inorganic nitrogen:dissolved inorganic phosphorous (din_dip), total dissolved nitrogen (tdn), dissolved organic nitrogen (don), dissolved organic carbon (doc), total organic carbon (toc), total nitrogen (tn).

## `tx_master.csv`

A compilation of all cleaned variables mentioned above

## `dust_master.csv`

Manually created dust data examining different dust permutations: time points, sums, averages (helpful for statistics)

#### Variables

date, time at collection 1 hour (t1), 7 hour (t7), 13 hour (t13), and 19 hour (t19). Sum of all 4 time points (tsum), average of all four time points (tavg), 24 hour hour composite sum (t7sum), 24 hour composite average (t7sum).

### 2023-05-01 data/raw_data/rawdata_README.Rmd
editing for grammer and flow
Updated README and citations

This folder contains dust concentrations in AOT from the Naval Reserach lab, environmental parameters (temperature and salinity) that were collected during the summer at the three sites, and  qPCR results for Total Vibrio.

>>>>

This folder contains two raw data sets:

## `dust.csv`

#### Variables:

yyymmddtt: Year_month_day_time of collection

dust: dust aerosol optical density. Concentrations (AOT) were collected at 4 time points (UTC): T0,T6,T12, and T18.

smoke: smoke

seasalt: seasalt

SS Total AOT: Total aerosol optical thickness

date: date

type: denotes daily sampling time period

## `copies_master.csv`

#### Variables:

Type: Denoted daily sampling period

Sample: Site and Day short code. Ex. BO11 = Blind Oso on 7/11/23

Cq: Cycle threshold value from qPCR. Indicated the cycle number that amplification was detected. A lower Cq value indicates earlier detection, and a higher concentration of target,\

### 2023-05-01 products/manuscript/Manuscript.qmd
Updated README and citations

---
title: "Nathan Greenslit Data Analysis Project"
subtitle: ""
author: Nathan Greenslit
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: false
    highlight-style: github
bibliography: MADA.bib
---

@@

*Vibrio* gene copies/mL was found to significantly correlate with dust aerosol optical density at Blind Oso (r = 0.7, p = 0.01), Canals (r = 0.75, p = 0.005), and the Gulf (r = 0.67, p = 0.02). The introduction of dust constituents have the potential to serve as a temporary alleviation from the limitation of nutrients that are essential for bloom formation. *Vibrio* species have one of the fastest doubling times among bacteria [@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes [@ringgaardChemotaxisArraysVibrio2018; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016]. Initially comprising a small proportion of the marine microbial communities (\<1%), *Vibrio* can come to quickly dominate, and thus have been classified as 'oppurtunitrophs'. The following declines in *Vibrio* gene copies/mL may be attributed to top down control, such as viral lysis [@molina-quiroz2023], or bottom-up controls due to an exhaustion of the resources provided by the dust [@psenner1992].

@@

*Vibrio* growth was observed at Blind Oso Bay, but slightly dampened in comparison to the Gulf. This site is an enclosed shallow tributary that frequently experiences high levels of nutrients (e.g.total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP) [@wetz]. The Bay experienced a dramatic drop in salinity from July 11th to July 19th that could be attributed to discharge from the OSP (Figure 3). While growth was significantly correlated with dust deposition, there are many environmental factors at play at this site, limiting the identification of the specific mechanisms that are eliciting this response.

>>>>

---
title: "Dusty Days and a Vibrio Craze: The Influence of Saharan Dust Deposition on Microbial Communities in High Nutrient Coastal Waters "
subtitle: ""
author: Nathan Greenslit
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: false
    highlight-style: github
bibliography: MADA.bib
---

@@

*Vibrio* gene copies/mL was found to significantly correlate with dust aerosol optical density at Blind Oso (r = 0.7, p = 0.01), Canals (r = 0.75, p = 0.005), and the Gulf (r = 0.67, p = 0.02). The introduction of dust constituents have the potential to serve as a temporary alleviation from the limitation of nutrients that are essential for bloom formation. *Vibrio* species have one of the fastest doubling times among bacteria [@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes [@ringgaardChemotaxisArraysVibrio2018; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016]. Initially comprising a small proportion of the marine microbial communities (\<1%), *Vibrio* can come to quickly dominate, and thus have been classified as 'oppurtunitrophs'. The following declines in *Vibrio* gene copies/mL may be attributed to top down control, such as viral lysis [@molina-quirozInteractionsVibrioPhages2023], or bottom-up controls due to an exhaustion of the resources provided by the dust [@psennerAreRapidChanges1992].

@@

*Vibrio* growth was observed at Blind Oso Bay, but slightly dampened in comparison to the Gulf. This site is an enclosed shallow tributary that frequently experiences high levels of nutrients (e.g.total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP) (Wetz, 2014). The Bay experienced a dramatic drop in salinity from July 11th to July 19th that could be attributed to discharge from the OSP (Figure 3). While growth was significantly correlated with dust deposition, there are many environmental factors at play at this site, limiting the identification of the specific mechanisms that are eliciting this response.
